---
title: "Demographics"
author: "Rene Michel"
date: '2022-08-03'
output: html_document
---

# Configuration

### Report settings

```{r report settings}

# R Markdown report settings
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE, 
                      out.width = "100%")

# Export .RData files for subsequent analyses?
export = TRUE

``` 

### Define paths and dependencies

```{r cfg paths, comment = ""}
# PATHS -----------------------------------------------------------------------
repo_path <- "add_path_here"
# REQUIRED PKGS ----------------------------------------------------------------
git_pkgs <- c("remichel/rmTools@HEAD")
cran_pkgs <- c("dplyr", "tidyr", "ggplot2", "knitr", "kableExtra", "readxl")

```

## Load packages

```{r pks, comment = ""}
# load and install required packages
rmTools::rm_install_packages(
  git_pkgs = git_pkgs,
  git_private_pkgs = NULL,
  cran_pkgs = cran_pkgs, load_after_install = T
)
```
## Load data

```{r load data, comment = ""}
demo <- read_xlsx(paste0(repo_path, "demographics.xlsx"))
```

# Demographics

```{r descriptive stats, comment = ""}


demo %>%
  # filter excluded participants
  filter(!(ID %in% c(1,3,4,20,21))) %>%
  summary()

demo %>% 
# filter excluded participants
  filter(!(ID %in% c(1,3,4,20,21))) %>%
  summarise(Age_sd = sd(Age))

demo %>%
  # filter excluded participants
  filter(!(ID %in% c(1,3,4,20,21))) %>%
  count(Sex)

demo %>%
  # filter excluded participants
  filter(!(ID %in% c(1,3,4,20,21))) %>%
  count(Handedness)


demo %>%
  # filter excluded participants
  filter(!(ID %in% c(1,3,4,20,21))) %>%
  count(`Dominant Eye`)