---
title: "Preprocessing"
author: "Rene Michel"
date: '2022-06-09'
output: html_document
---

# Configuration

```{r report settings, message = F, warning= F}

# R Markdown report settings
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE, 
                      out.width = "100%")

# Export .RData files for subsequent analyses?
export = TRUE

``` 

```{r cfg paths, message = F, warning= F}
# PATHS -----------------------------------------------------------------------
repo_path <- "add_path_here"
# REQUIRED PKGS ----------------------------------------------------------------
git_pkgs <- c("remichel/rmTools@HEAD")
cran_pkgs <- c("dplyr", "tidyr", "ggplot2", "knitr", "kableExtra")

```

```{r pks, message = F, warning= F}
# load and install required packages
rmTools::rm_install_packages(
  git_pkgs = git_pkgs,
  git_private_pkgs = NULL,
  cran_pkgs = cran_pkgs, load_after_install = T
)
```
## Design specs and exclusion criteria


```{r cfg analysis, echo = T, message = F, warning= F}
# DESIGN SPECIFICATIONS -------------------------------------------------

# list of subject ids
subject_list <- seq(1, 44, 1)
# inter frame interval of our monitor (in ms)
ifi <- 1000 / 120
# spacing of SOAs (in number of frames)
soa_spacing <- 5
# expected SOAs (in ms)
expected_soas <- round(c(0, seq(5, 125, soa_spacing) * ifi), 1)
# values in dataframe that should be interpreted as "NA"
na_strings <- c("9999", "9999.00")

# EXCLUSION CRITERIA ----------------------------------------------------

# subjects with high FA rate in catch trials will be excluded; .2 = 20%
fa_threshold <- .2
# RTs lower than min_rt will be excluded; in ms
min_rt <- 200
# exclusion criteria for RTs, e.g. values exceeding +/- 4*SD will be excluded
rt_sd <- 4

```


# Load datasets

```{r load data, message = F, warning= F}
# load behavioral data
for (i in subject_list) {
  if (i == min(subject_list)) {
    raw <- read.table(paste0(repo_path, i, "/", i, ".txt"),
      h = T,
      na.strings = na_strings
    )
  } else {
    # concatenate
    raw <- rbind(raw, read.table(paste0(repo_path, i, "/", i, ".txt"),
      h = T,
      na.strings = na_strings
    ))
  }
}
# load outofbounds file, which stores information about all trials that n
# eeded to be aborted online but were missed by online fixation control
load(paste0(repo_path, "outofbounds.Rdat"))

```

# Preparation


```{r preparation, message = F, warning= F}

# align SOAs across subjects
raw <- raw %>%
  rowwise() %>%
  mutate(soa = expected_soas[which.min((abs(soa - expected_soas)))]) %>%
  ungroup()


# introduce validity variable
raw <- raw %>%
  mutate(validity = case_when(
    target_position == 0 ~ "catch",
    cue_position == target_position ~ "valid",
    TRUE ~ "invalid"
  ))

# prepare outofbounds df
outofbounds <- outofbounds %>%
  mutate(needs_abortion = 1) %>%
  mutate(subject = as.numeric(subject))
```





# Exclusions

We apply the following exclusions:

- subjects with a false alarm rate > 20% (note that these subjects were subsequently substituted by new recorded subjects to end with the desired sample size of 39 subjects)
- practice trials (not needed for analysis)
- aborted trials
- trials that were missed by online fixation control
- catch trials (not needed for analysis, only used to determine subjects false alarm rates)
- incorrect trials (i.e. no button press)
- trials with an RT < 200ms
- trials with an RT exceeding mean(RT) +/- 4 SD (within subject and within conditions, after excluding all abovementioned trials)


## False alarms

The following subjects need to be excluded for subsequent analyses due to a false alarm rate higher than 20%:

```{r fa exclusions, message = F, warning= F}

fa <- raw %>%
  dplyr::filter(section == 1, 
                soa == 0, 
                abort == 0) %>%
  dplyr::group_by(subject) %>%
  dplyr::summarise(n_fa = n() - sum(correct),
                   ntrials = n(),
                   fa = round(1 - mean(correct), 3)) %>%
  dplyr::mutate(fa_threshold = fa_threshold,
                exclude = case_when(fa_threshold < fa ~ 1, TRUE ~ 0)) %>%
  arrange(desc(n_fa)) %>%
  as.data.frame() 

# table excluded subjects
fa %>%
  filter(exclude == TRUE) %>%
  knitr::kable(.) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>% 
  scroll_box(width = "100%", height = "300px")


```


```{r exclusion, message = F, warning= F}

clean <- raw %>%
  # exclude practice trials & aborted trials
  filter(section == 1, abort == 0) %>%
  # exclude trials that were missed to be aborted and repeated online
  full_join(., outofbounds) %>%
  filter(is.na(needs_abortion)) %>%
  # exclude subjects with too high FA rate
  filter(!subject %in% filter(fa, exclude == 1)$subject) %>%
  # exclude catch trials and incorrect trials
  filter(soa != 0, correct == 1) %>%
  # exclude RTs <= 200ms
  filter(rt >= min_rt) %>%
  # prepare lower and upper bound per subject/condition to exlude RTs > 4 * SD
  # in a subsequent step
  group_by(subject, validity) %>%
  mutate(lower = case_when(mean(rt) - rt_sd * sd(rt) > 
                             min_rt ~ mean(rt) - rt_sd * sd(rt),
                           TRUE ~ min_rt),
         upper = mean(rt) + rt_sd * sd(rt)) %>%
  # now, exclude RTs > 4 * SD (per participant & validity)
  filter(rt < upper & rt > lower) %>%
  ungroup() %>%
  # tidy up dataframe by removing irrelevant variables for subsequent analyses
  select(-c(response, abort, section, correct, cue_position, target_position))


```

## Missed rejections

Number of trials excluded due to missed rejections.

```{r missed rej, message = F, warning= F}

raw %>%
  # exclude practice trials & aborted trials, also exclude catch trials and incorrect trials
  filter(section == 1, abort == 0, soa != 0, correct == 1) %>%
          # join with out of bounds df
  left_join(., outofbounds, by = c("subject", "trial")) %>%
  # exclude subjects with too high FA rate
  filter(!subject %in% filter(fa, exclude == 1)$subject) %>%
  group_by(subject) %>%
  summarise(needs_abortion = sum(needs_abortion, na.rm = T)) %>%
  ungroup() %>%
  summarise(min = min(needs_abortion),
            mean = mean(needs_abortion),
            max = max(needs_abortion),
            sd = sd(needs_abortion)) %>%
  knitr::kable(.) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))



```


## Exlusions due to RT boundaries

Number of trials excluded due to RT boundaries.

```{r excl due to rt, message = F, warning= F}


raw %>%
  # exclude practice trials & aborted trials
  filter(section == 1, abort == 0) %>%
  # exclude trials that were missed to be aborted and repeated online
  full_join(., outofbounds) %>%
  filter(is.na(needs_abortion)) %>%
  # exclude subjects with too high FA rate
  filter(!subject %in% filter(fa, exclude == 1)$subject) %>%
  # exclude catch trials and incorrect trials
  filter(soa != 0, correct == 1) %>%
  # prepare lower and upper bound per subject/condition to exlude RTs > 4 * SD
  # in a subsequent step
  group_by(subject, validity) %>%
  mutate(lower = case_when(mean(rt) - rt_sd * sd(rt) > 
                             min_rt ~ mean(rt) - rt_sd * sd(rt),
                           TRUE ~ min_rt),
         upper = mean(rt) + rt_sd * sd(rt)) %>%
  # now, exclude RTs > 4 * SD (per participant & validity)
  filter(rt > upper | (rt < lower | rt < 200)) %>%
  ungroup() %>%
  count(subject) %>%
  summarize(min_n = min(n),
            mean_n = mean(n),
            max_n = max(n),
            sd_n = sd(n)) %>%
  knitr::kable(.) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))


```


## Data loss 

Percentage of data loss.

```{r data loss, message = F, warning= F}
clean %>%
  count(subject) %>%
  mutate(p_loss = 1-(n/600)) %>%
  summarize(min_p_loss = min(p_loss),
            mean_p_loss = mean(p_loss),
            max_p_loss = max(p_loss),
            sd_p_loss = sd(p_loss))  %>%
  knitr::kable(.) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## Remaining trials per cell

How many trials per SOA/validity combination are left after applying the above-mentioned exclusions?


```{r tpc condition, message = F, warning= F}

clean %>% 
  group_by(subject, validity, soa) %>% 
  summarize(n = n()) %>%
  summarize(min = min(.data$n),
            mean = mean(.data$n),
            max = max(.data$n)) %>%
  ungroup() %>%
  summarize(min = min(.data$min),
            mean = mean(.data$mean),
            max = max(.data$max)) %>%
  knitr::kable(.) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>% 
  scroll_box(width = "100%", height = "90px")



```




# Z-Scaling

RTs were z-scaled within subjects regardless of SOA and validity.

```{r scaling, message = F, warning= F}

# z-scale RTs
clean <- clean %>%
  group_by(subject) %>%
  mutate(rtz = as.numeric(scale(rt))) %>%
  select(-c(lower, upper, needs_abortion)) %>%
  relocate(validity, .before = rt) 

```

# Difference Time Courses

Creating difference time courses by computing the difference "invalid-valid" on the single subject level. The single subject time courses are then averaged across subjects.

```{r difference time course ss, , message = F, warning= F, fig.height=unit(8, "cm")}

# single subject difference time courses
diff <- clean %>%
  group_by(subject, validity, soa) %>%
  summarize(rt = mean(rt),
            rtz = mean(rtz)) %>%
  pivot_wider(names_from = validity,
              values_from = c("rt", "rtz")) %>%
  rowwise() %>%
  mutate(diff = rt_invalid - rt_valid,
         diffz = rtz_invalid - rtz_valid) %>%
  ungroup() %>%
  select(-c(rt_invalid, rt_valid, rtz_invalid, rtz_valid))

# grand average difference time course
# compute grand average difference time course
diff_agg <- diff %>%
  group_by(soa) %>%
  summarize(diff = mean(diffz)) 


```



# Export

Exporting preprocessed datasets (raw, cleaned, single subject difference time courses, grand average difference time course)

```{r export}
if(export == TRUE){
  save(file = paste0(repo_path, "raw.RData"), raw)
  save(file = paste0(repo_path, "clean.RData"), clean)
  save(file = paste0(repo_path, "diff.RData"), diff)
  save(file = paste0(repo_path, "diff_agg.RData"), diff_agg)
}

```