---
title: "Modelling Analysis"
author: "Rene Michel"
date: '2022-06-09'
output: html_document
---


# Configuration

### Report settings

```{r report settings, message = FALSE, warning=FALSE}

knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE, 
                      out.width = "100%")

do_save = TRUE

``` 

### Define paths and dependencies

```{r cfg paths, message = FALSE, warning=FALSE}
# PATHS -----------------------------------------------------------------------
repo_path <- "add_path_here"
# REQUIRED PKGS ----------------------------------------------------------------
git_pkgs <- c("remichel/rmTools@HEAD")
cran_pkgs <- c("Rmisc", "dplyr", "tidyr", "tibble", "nls.multstart", "broom",
               "AICcmodavg", "ggplot2", "knitr", "kableExtra", "wesanderson")
```

```{r pks, message = FALSE, warning=FALSE}
# load and install required packages
rmTools::rm_install_packages(
  git_pkgs = git_pkgs,
  git_private_pkgs = NULL,
  cran_pkgs = cran_pkgs, load_after_install = T
)
```

### Define analysis parameters

Set all required parameters for the subsequent analysis. 


```{r cfg analysis, echo = T, message = FALSE, warning=FALSE}

# number of start values for fitting procedure of nls_multstart
# will be repeated times the number of parameters to be fitted,
# e.g. rep(niter, 3) for three parameters (see "? nls_multstart")
niter <- 5
# no early stopping rule for fitting procedure with nls_multstart,
# setting convergence count to FALSE forces to do all iterations
convergence_count <- FALSE
# parameter space limits per parameter (lower and upper bounds)
# Note 1: for f, we exclude nyquist frequency (i.e. 12 Hz)
# Note 2: lowest technically plausible frequency is 1 Hz
limits <- list(
  i = c(-Inf, Inf),
  n0 = c(0, Inf),
  tau = c(0, 3),
  a = c(0, Inf),
  f = c(1, 11), # 3-8
  phi = c(0, 2 * pi)
)
# start value range per parameter (lower and upper bounds)
# for grid search, equally spaced start values will be drawn from
# the start value range
starts <- list(
  i = c(-5, 5),
  n0 = c(0, 5),
  tau = c(0, 3),
  a = c(0, 5),
  f = c(1, 11), # 3-8
  phi = c(0, 2 * pi)
)

# MODEL FUNCTIONS -------------------------------------------------------

# exponential model formula
expModel <- function(t, intercept, n0, tau) {
  intercept + n0 * exp(-t / tau)
}
# sinusoidal model formula
sinModel <- function(t, intercept, a, f, phi) {
  intercept + a * sin(2 * pi * t * f + phi)
}
# hybrid model formula
hybModel <- function(t, intercept, n0, tau, a, f, phi) {
  intercept + n0 * exp(-t / tau) + a * sin(2 * pi * t * f + phi)
}

```


# Load Data

```{r load, message = FALSE, warning=FALSE}
load(paste0(repo_path, "diff.RData"))
load(paste0(repo_path, "diff_agg.RData"))
```

# Prepare datasets

We need SOA in s, not ms, to interpret frequency estimates properly. 

```{r prepare datasets, message = FALSE, warning=FALSE}

modeldata_ga <- diff_agg %>%
  rename(diffz = diff) %>%
  mutate(soa = soa/1000) %>% 
  # we need this variable later for joining with the model results and predictions
  mutate(model = "observed") 


modeldata_ss <- diff %>%
  mutate(soa = soa/1000) %>%
  # we need this variable later for joining with the model results and predictions
  mutate(model = "observed") 

```


# Model comparison


### Grand Average



```{r modelling grandavg, message = FALSE, warning=FALSE}

# set seed to get reproducible results
set.seed(1)

# note for modelling: 
# some iterations will throw errors due to bad starting values, we can
# suppress such with supp_errors. If all iterations fail, it will still throw
# an error.

results <- modeldata_ga %>%
  # nest data to allow for modelling within tibble
  nest(data = everything()) %>%
  # fit all models
  mutate(m_int = purrr::map(.data$data, ~ lm(diffz ~ 1, data = .))) %>%
  mutate(m_lin = purrr::map(.data$data, ~ lm(diffz ~ soa, data = .))) %>%
  mutate(m_exp = 
           purrr::map(.data$data, 
                      ~ nls_multstart(diffz ~ expModel(soa, intercept, n0, tau),
                                      data = .,
                                      lower = c(intercept = limits$i[1],
                                                n0 = limits$n0[1], 
                                                tau = limits$tau[1]),
                                      upper = c(intercept = limits$i[2],
                                                n0 = limits$n0[2], 
                                                tau = limits$tau[2]),
                                      start_lower = 
                                        c(intercept = starts$i[1], 
                                          n0 = starts$n0[1], 
                                          tau = starts$tau[1]),
                                      start_upper = 
                                        c(intercept = starts$i[2], 
                                          n0 = starts$n0[2], 
                                          tau = starts$tau[2]),
                                      iter = rep(niter, 3),
                                      supp_errors = "Y", 
                                      convergence_count = convergence_count
                                      ))) %>%
  mutate(m_sin = 
           purrr::map(.data$data, 
                      ~ nls_multstart(diffz ~ sinModel(soa, 
                                                     intercept, a, f, phi),
                                      data = .,
                                      lower = c(intercept = limits$i[1], 
                                                a = limits$a[1], 
                                                f = limits$f[1], 
                                                phi = limits$phi[1]),
                                      upper = c(intercept = limits$i[2], 
                                                a = limits$a[2], 
                                                f = limits$f[2], 
                                                phi = limits$phi[2]),
                                      start_lower = c(intercept = starts$i[1],
                                                      a = starts$a[1], 
                                                      f = starts$f[1], 
                                                      phi = starts$phi[1]),
                                      start_upper = c(intercept = starts$i[2],
                                                      a = starts$a[2], 
                                                      f = starts$f[1], 
                                                      phi = starts$phi[2]),
                                      iter = rep(niter, 4),
                                      supp_errors = "Y",
                                      convergence_count = convergence_count))) %>%
  mutate(m_hyb = 
           purrr::map(.data$data, 
                      ~ nls_multstart(diffz ~ hybModel(soa, intercept, 
                                                     n0, tau, a, f, phi),
                                      data = .,
                                      lower = c(intercept = limits$i[1], 
                                                n0 = limits$n0[1], 
                                                tau = limits$tau[1], 
                                                a = limits$a[1], 
                                                f = limits$f[1], 
                                                phi = limits$phi[1]),
                                      upper = c(intercept = limits$i[2], 
                                                n0 = limits$n0[2], 
                                                tau = limits$tau[2], 
                                                a = limits$a[2], 
                                                f = limits$f[2], 
                                                phi = limits$phi[2]),
                                      start_lower = c(intercept = starts$i[1],
                                                      n0 = starts$n0[1], 
                                                      tau = starts$tau[1], 
                                                      a = starts$a[1], 
                                                      f = starts$f[1], 
                                                      phi = starts$phi[1]),
                                      start_upper = c(intercept = starts$i[2],
                                                      n0 = starts$n0[2], 
                                                      tau = starts$tau[2], 
                                                      a = starts$a[2], 
                                                      f = starts$f[2], 
                                                      phi = starts$phi[2]),
                                      iter = rep(niter, 6),
                                      supp_errors = "Y", 
                                      convergence_count = convergence_count))) %>%
  # bring into long format
  pivot_longer(cols = c(m_int, m_lin, m_exp, m_sin, m_hyb),
               names_to = "model",
               values_to = "fit") %>%
  # extract goodness of fits per model
  group_by(fit) %>%
  mutate(glance = purrr::map(.data$fit, ~ broom::glance(.))) %>%
  # extract param estimates per model
  mutate(tidy = purrr::map(.data$fit, ~ broom::tidy(.))) %>%
  # unpack results
  unnest(cols = c(glance)) %>%
  # remove unwanted cols
  select(-c(sigma, isConv, finTol, logLik, 
            deviance, statistic, df, p.value)) %>%
  # unpack estimates
  unnest(cols = c(tidy)) %>%
  # remove unwanted cols
  select(-c(std.error, statistic, p.value)) %>%
  # align parameter names (they differ between stats and nls_multstart pkgs)
  mutate(term = case_when(term == "(Intercept)" ~ "intercept",
                          TRUE ~ term)) %>%
  # bring estimates into wide format
  pivot_wider(names_from = term,
              values_from = estimate) %>%
  # calculate AICc
  mutate(AICc = purrr::map(.data$fit, ~ AICc(.))) %>%
  unnest(AICc) %>%
  # get model predictions for diffz
  mutate(diffz = purrr::map(.data$fit, ~ predict(.))) %>%
  unnest(cols = diffz) %>%
  # rename soa to b1, because linear model's b1 was called "soa"
  rename(b1 = soa) %>% 
  # add SOA as variable to join with modeldata_ga in a next step
  mutate(soa = sort(unique(modeldata_ga$soa))) %>%
  # add observed values
  select(-data) %>%
  full_join(modeldata_ga) %>%
  # convert GOF indices to long format plotting
  pivot_longer(cols = c(AIC, AICc, BIC),
               names_to = "index",
               values_to = "index_value")

```

##### Plot model predictions

```{r plot predictions, message = FALSE, warning=FALSE}



results <- results %>%
  mutate(soa = soa * 1000) %>%
  filter(model != "observed")


pdat <- diff %>%
  summarySEwithin(data = ., 
                  measurevar = "diffz",
                  withinvars = c("soa"),
                  idvar = "subject") %>%
  mutate(soa = fac2num(soa),
         observed = diffz) %>%
  select(-c("diffz")) %>%
  full_join(results, by = c("soa")) %>%
  mutate(model = factor(model, 
                        levels = c("m_int", "m_lin", "m_exp", "m_sin", "m_hyb"), 
                        labels = c("intercept", "linear", "exponential", "rhythmic", "hybrid")))

labels = rep('', 25)
labels[seq(1,25,7)] = as.character(sort(unique(pdat$soa))[seq(1,25,7)])

model_plot <- pdat %>%
  ggplot(aes(x = soa, y = diffz)) +
  geom_line(aes(x = soa, y = diffz, color = model), size = 3)+
  geom_hline(yintercept = 0, color = "grey50", linetype = "dotted", size = 1.2)+
  geom_point(aes(x = soa, y = observed), size = 2.3, position=position_dodge(width=8))+
  labs(x = "CTOA", y = "z-standardized RT difference\n(invalid-valid)")+
  facet_wrap(~model, ncol = 5)+
  scale_x_continuous(expand = c(0.02, 0), breaks = sort(unique(pdat$soa)), labels = labels) +
  scale_y_continuous(expand = c(0.05, 0)) +
  scale_color_manual(values = wes_palette("Darjeeling1"))+
  theme_bw() +
  theme(legend.position="none",
        axis.title = element_text(face = "bold", size = 16),
        axis.text.y = element_text(face = "bold", size = 12),
        axis.text.x = element_text(face = "bold", size = 9),
        legend.title = element_blank(),
        legend.text = element_text(face = "bold", size = 12),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold", size = 15),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank())

model_plot



if(do_save) ggsave(filename = paste0(repo_path, "model_plot_", paste0(starts$f, collapse = "_"), ".svg"), plot = model_plot, width = 3600, height = 1200, units = "px")


```

##### Model fits

```{r Fits, message = FALSE, warning=FALSE}

# show table with results and fitted params
result_table <- results %>%
  ungroup() %>%
  filter(model != "observed") %>%
  select(-c(r.squared, adj.r.squared, fit, soa, diffz, nobs)) %>%
  distinct() %>%
  pivot_wider(names_from = index,
              values_from = index_value)

result_table %>%
  knitr::kable(.) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  scroll_box(width = "100%", height = "300px")

# latex table
result_table %>%
  select(-c("df.residual", "AIC")) %>%
  knitr::kable(., "latex", align = "c", digits = 2) 



```

##### Model comparisons using AICc and BIC


```{r AICc and BIC, message = FALSE, warning=FALSE}

dAICc <- result_table %>%
  # AICc differences from best model, and relative likelihood of each model
  mutate(minAICc = min(AICc),
         dAICc = AICc - minAICc,
         relL = exp(-0.5 * dAICc)) %>%
  # compute Akaike weights
  mutate(w_i = relL / sum(relL)) %>%
  # Akaike weight ratios
  bind_cols(., outer(.$w_i, .$w_i, "/") %>% 
  as.data.frame()  %>% 
  setNames(., nm = c("vs_intercept", "vs_lin", "vs_exp", "vs_sin", "vs_hyb"))) %>%
  pivot_longer(cols = c("vs_intercept", "vs_lin", "vs_exp", "vs_sin", "vs_hyb"),
               names_to = "vs",
               values_to = "w_ratio")  


dBIC <- result_table %>%
  # add all BIC differences between models
  bind_cols(., outer(.$BIC, .$BIC, "-") %>% 
  as.data.frame()  %>% 
  setNames(., nm = c("vs_intercept", "vs_lin", "vs_exp", "vs_sin", "vs_hyb"))) %>%
  pivot_longer(cols = c("vs_intercept", "vs_lin", "vs_exp", "vs_sin", "vs_hyb"),
               names_to = "vs",
               values_to = "dBIC") %>%
  # add BF approximation by Wagenmakers (2007)
  mutate(approxBF = exp(.5 * dBIC)) 

# combine both tables
dAICc %>% 
  full_join(dBIC) %>%
  select(-c(df.residual, intercept, b1, n0, tau, a, f, phi, AIC)) %>%
  knitr::kable(., digits = 3, format.args = list(scientific = FALSE)) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))    %>%
  scroll_box(width = "100%", height = "300px")





```

### Single subject


```{r modelling ssubj, message = FALSE, warning=FALSE}

# set seed to get reproducible results
set.seed(999)


results <- modeldata_ss %>%
  group_by(subject) %>%
  # nest data to allow for modelling within tibble
  nest() %>%
  # fit all models
  mutate(m_int = purrr::map(.data$data, ~ lm(diffz ~ 1, data = .))) %>%
  mutate(m_lin = purrr::map(.data$data, ~ lm(diffz ~ soa, data = .))) %>%
  mutate(m_exp = 
           purrr::map(.data$data, 
                      ~ nls_multstart(diffz ~ expModel(soa, intercept, n0, tau),
                                      data = .,
                                      lower = c(intercept = limits$i[1], 
                                                n0 = limits$n0[1], 
                                                tau = limits$tau[1]),
                                      upper = c(intercept = limits$i[2], 
                                                n0 = limits$n0[2], 
                                                tau = limits$tau[2]),
                                      start_lower = c(intercept = starts$i[1],
                                                      n0 = starts$n0[1], 
                                                      tau = starts$tau[1]),
                                      start_upper = c(intercept = starts$i[2],
                                                      n0 = starts$n0[2], 
                                                      tau = starts$tau[2]),
                                      iter = rep(niter, 3),
                                      supp_errors = "Y", 
                                      convergence_count = convergence_count))) %>%
  mutate(m_sin = 
           purrr::map(.data$data, 
                      ~ nls_multstart(diffz ~ sinModel(soa, intercept, a, 
                                                     f, phi),
                                      data = .,
                                      lower = c(intercept = limits$i[1], 
                                                a = limits$a[1], 
                                                f = limits$f[1], 
                                                phi = limits$phi[1]),
                                      upper = c(intercept = limits$i[2], 
                                                a = limits$a[2], 
                                                f = limits$f[2], 
                                                phi = limits$phi[2]),
                                      start_lower = c(intercept = starts$i[1],
                                                      a = starts$a[1], 
                                                      f = starts$f[1], 
                                                      phi = starts$phi[1]),
                                      start_upper = c(intercept = starts$i[2],
                                                      a = starts$a[2], 
                                                      f = starts$f[1], 
                                                      phi = starts$phi[2]),
                                      iter = rep(niter, 4),
                                      supp_errors = "Y",
                                      convergence_count = convergence_count))) %>%
  mutate(m_hyb = 
           purrr::map(.data$data, 
                      ~ nls_multstart(diffz ~ hybModel(soa, intercept, 
                                                     n0, tau, a, f, phi),
                                      data = .,
                                      lower = c(intercept = limits$i[1], 
                                                n0 = limits$n0[1], 
                                                tau = limits$tau[1], 
                                                a = limits$a[1], 
                                                f = limits$f[1], 
                                                phi = limits$phi[1]),
                                      upper = c(intercept = limits$i[2], 
                                                n0 = limits$n0[2], 
                                                tau = limits$tau[2], 
                                                a = limits$a[2], 
                                                f = limits$f[2], 
                                                phi = limits$phi[2]),
                                      start_lower = c(intercept = starts$i[1],
                                                      n0 = starts$n0[1], 
                                                      tau = starts$tau[1], 
                                                      a = starts$a[1], 
                                                      f = starts$f[1], 
                                                      phi = starts$phi[1]),
                                      start_upper = c(intercept = starts$i[2],
                                                      n0 = starts$n0[2], 
                                                      tau = starts$tau[2], 
                                                      a = starts$a[2], 
                                                      f = starts$f[2],
                                                      phi = starts$phi[2]),
                                      iter = rep(niter, 6),
                                      supp_errors = "Y",
                                      convergence_count = convergence_count))) %>%
  # bring into long format
  pivot_longer(cols = c(m_int, m_lin, m_exp, m_sin, m_hyb),
               names_to = "model",
               values_to = "fit") %>%
  # extract goodness of fits per model
  group_by(subject, fit) %>%
  mutate(glance = purrr::map(.data$fit, ~ broom::glance(.))) %>%
  # extract param estimates per model
  mutate(tidy = purrr::map(.data$fit, ~ broom::tidy(.))) %>%
  # unpack results
  unnest(cols = c(glance)) %>%
  # remove unwanted cols
  select(-c(sigma, isConv, finTol, logLik, 
            deviance, statistic, df, p.value)) %>%
  # unpack estimates
  unnest(cols = c(tidy)) %>%
  # remove unwanted cols
  select(-c(std.error, statistic, p.value)) %>%
  # align parameter names (they differ between stats and nls_multstart pkgs)
  mutate(term = case_when(term == "(Intercept)" ~ "intercept",
                          TRUE ~ term)) %>%
  # bring estimates into wide format
  pivot_wider(names_from = term,
              values_from = estimate) %>%
  # calculate AICc
  mutate(AICc = purrr::map(.data$fit, ~ AICc(.))) %>%
  unnest(AICc) %>%
  # get model predictions for diffz
  mutate(diffz = purrr::map(.data$fit, ~ predict(.))) %>%
  unnest(cols = diffz) %>%
  # rename soa to b1, because linear model's b1 was called "soa"
  rename(b1 = soa) %>% 
  # add SOA as variable to join with modeldata_ss in a next step
  mutate(soa = sort(unique(modeldata_ss$soa))) %>%
  # add observed values
  select(-data) %>%
  full_join(modeldata_ss) %>%
  # convert GOF indices to long format plotting
  pivot_longer(cols = c(AIC, AICc, BIC),
               names_to = "index",
               values_to = "index_value") 


```


```{r identify winner model per subj, message = FALSE, warning=FALSE}

results <- results %>%
  # mark winner model per subject and index
  group_by(subject, index) %>%
  mutate(min_value = min(index_value, na.rm = T)) %>% # na.rm = T to ignore "observed" column
  mutate(winner_model = case_when(index_value == min_value ~ "*",
                                TRUE ~ "")) %>%
  ungroup()

```

##### Model fits

```{r ss model fits, message = FALSE, warning=FALSE}

# show table with results and fitted params
result_table <- results %>% 
  ungroup() %>%
  filter(model != "observed") %>%
  mutate(n0 = round(n0,2),
         tau = round(tau, 2),
         a = round(a, 2),
         f = round(f, 2),
         phi = round(phi, 2)) %>%
  select(-c(r.squared, adj.r.squared, fit, soa, diffz, min_value, diff, nobs, df.residual, winner_model)) %>%
  distinct() %>%
  pivot_wider(names_from = index,
              values_from = index_value) 

result_table %>%
  knitr::kable(.) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  scroll_box(width = "100%", height = "300px")

```



##### Model comparison using AICc and BIC

```{r ss AIC BIC, message = FALSE, warning=FALSE}

dAICc <- result_table %>%
  group_by(subject) %>%
  # AICc differences from best model, and relative likelihood of each model
  mutate(minAICc = min(AICc),
         dAICc = AICc - minAICc,
         relL = exp(-0.5 * dAICc)) %>%
  # compute Akaike weights
  mutate(w_i = relL / sum(relL)) %>%
  group_modify( ~ bind_cols(., outer(.$w_i, .$w_i, "/") %>% as.data.frame() %>% 
               setNames(., nm = c("vs_intercept", "vs_lin", "vs_exp", "vs_sin", "vs_hyb")))) %>% 
  group_by(subject) %>%
  pivot_longer(cols = c("vs_intercept", "vs_lin", "vs_exp", "vs_sin", "vs_hyb"),
               names_to = "vs",
               values_to = "w_ratio") 


dBIC <- result_table %>%
  # add all BIC differences between models for each subject
  group_by(subject) %>%
  group_modify( ~ bind_cols(., outer(.$BIC, .$BIC, "-") %>% as.data.frame() %>% 
               setNames(., nm = c("vs_intercept", "vs_lin", "vs_exp", "vs_sin", "vs_hyb")))) %>% 
  group_by(subject) %>%
  pivot_longer(cols = c("vs_intercept", "vs_lin", "vs_exp", "vs_sin", "vs_hyb"),
               names_to = "vs",
               values_to = "dBIC")  %>%
  # add BF approximation by Wagenmakers (2007)
  mutate(approxBF = exp(.5 * dBIC))

# combine both tables
deltas <- dAICc %>% 
  full_join(dBIC) %>%
  select(-c(intercept, b1, n0, tau, a, f, phi, AIC)) 

deltas %>%
  knitr::kable(.) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  scroll_box(width = "100%", height = "300px")

```


#### Distribution of winner models

```{r winner models, message = FALSE, warning=FALSE}

# distribution of winner models across subjects

winner_per_index <- results %>%
  filter(model != "observed", index != "AIC") %>%
  # convert to factor so that missing factor levels (e.g. because of 0 counts) can be added to final table
  purrr::modify_if(is.character, as.factor) %>%
  # only select necessary variables
  select(subject, model, index, winner_model) %>%
  # there were 25 rows per subject due to model predicitions per SOA, but we need to get rid of them here to count occurences of winner models
  distinct() %>%
  # helper variable to be counted in the next step
  group_by(index, model, winner_model) %>%
  tally() %>%
  ungroup() %>%
  # fill all combinations of the grouping variables that werent occuting in dataset with 0
  complete(., index, model, winner_model, fill = list(n = 0)) %>%
  # filter to winner models
  filter(winner_model == "*") %>%
  mutate(model = factor(model, 
                        levels = c("m_int", "m_lin", "m_exp", "m_sin", "m_hyb"), 
                        labels = c("intercept", "linear", "exponential", "sinusoidal", "hybrid")))



# table for distribution
winner_per_index %>%
  # exclude winnermodel var
  select(-c(winner_model)) %>%
  # bring into wide format
  pivot_wider(names_from = index, values_from = n) %>%
  # order chart with increasing complexity of model
  slice(3,4,1,5,2) %>%
  # fancy table
  knitr::kable(.) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  scroll_box(width = "100%", height = "300px")



```

