---
title: "IOR Analysis"
author: "Rene Michel"
date: '2022-06-09'
output: html_document
---

# Configuration

### Report settings

```{r report settings, message = FALSE, warning=FALSE}

knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE, 
                      out.width = "100%")

do_save = TRUE

``` 

### Define paths and dependencies

```{r cfg paths, message = FALSE, warning=FALSE}
# PATHS -----------------------------------------------------------------------
repo_path <- "insert_path_here"
# REQUIRED PKGS ----------------------------------------------------------------
git_pkgs <- c("remichel/rmTools@HEAD")
cran_pkgs <- c(
  "Rmisc", "dplyr", "tidyr", "afex", "ggplot2", "emmeans",
  "knitr", "kableExtra", "apa"
  )

```

```{r pks, message = FALSE, warning=FALSE}
# load and install required packages
rmTools::rm_install_packages(
  git_pkgs = git_pkgs,
  git_private_pkgs = NULL,
  cran_pkgs = cran_pkgs, load_after_install = T
)
```


### Define analysis parameters

Set all required parameters for the subsequent analysis. 


```{r cfg analysis, echo = T}
# SOA selection for ANOVA (based on Posner&Cohen 1981)
anova_soas <- c(41.7, 83.3, 208.3, 291.7, 500)
# multiple comparison correction for post hoc tests
mcc_method <- "bon"
```


# Load datasets

```{r load data, message = FALSE, warning=FALSE}
load(paste0(repo_path, "clean.Rdata"))
load(paste0(repo_path, "diff.Rdata"))
load(paste0(repo_path, "diff_agg.Rdata"))
```

# Descriptive Statistics (after exclusions)



## RTs

#### RTs by validity

```{r rt descr val, message = FALSE, warning=FALSE}

clean %>%
  group_by(subject, validity) %>%
  summarize(rt = mean(rt)) %>%
  group_by(validity) %>%
  summarize(mean_rt = mean(rt),
            sd_rt = sd(rt)) %>%
  knitr::kable(.) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```


## Grand average time course by validity

Errorbars represent Cosineau-corrected SE.

```{r ga tc, message = FALSE, warning=FALSE}


# RTZ publication plot
pdat <- clean %>%
  summarySEwithin(data = ., 
                  measurevar = "rtz",
                  withinvars = c("validity", "soa"),
                  idvar = "subject") %>%
  
  mutate(soa = fac2num(soa)) %>%
  mutate(highlight_dots = case_when(soa %in% anova_soas ~ 1,
                                    TRUE ~ 0))

labels = rep('', 25)
labels[seq(1,25,2)] = as.character(sort(unique(pdat$soa))[seq(1,25,2)])


tc_plot <- pdat %>%
  ggplot(aes(x = soa, y = rtz, group = validity)) +
  geom_hline(yintercept = 0, color = "grey50", linetype = "dotted", size = 1.2)+
  geom_line(aes(color = validity), size = 1.2, alpha = .5)+
  geom_linerange(aes(min = rtz - se, 
                    max = rtz + se, 
                    color = validity), size = 1.1,
                position=position_dodge(width=8))+
  geom_point(aes(color = validity), size = 2.3, position=position_dodge(width=8))+
  geom_point(data = pdat[pdat$highlight_dots == 1,], aes(fill = validity), size = 4, color = "black", shape = 21, position=position_dodge(width=8))+
  annotate("rect", xmin = anova_soas[1] - 18, xmax = anova_soas[1] + 18, ymin = .5, ymax = 1.1,
             alpha = 0, color= "black")+ 
  annotate("text", x = anova_soas[1], y = 1.12, label = "*", size = 10)+ 
  annotate("rect", xmin = anova_soas[2] - 18, xmax = anova_soas[2] + 18, ymin = .25, ymax = .6,
             alpha = 0, color= "black")+ 
  annotate("text", x = anova_soas[2], y = .635, label = "n.s.", size = 4, fontface = 2)+ 
  annotate("rect", xmin = anova_soas[3] - 18, xmax = anova_soas[3] + 18, ymin = 0, ymax = .25,
             alpha = 0, color= "black")+ 
  annotate("text", x = anova_soas[3], y = .285, label = "n.s.", size = 4, fontface = 2)+ 
  annotate("rect", xmin = anova_soas[4] - 18, xmax = anova_soas[4] + 18, ymin = -.35, ymax = .1,
             alpha = 0, color= "black")+
  annotate("text", x = anova_soas[4], y = .12, label = "*", size = 10)+ 
  annotate("rect", xmin = anova_soas[5] - 18, xmax = anova_soas[5] + 18, ymin = -.6, ymax = .2,
             alpha = 0, color= "black")+
  annotate("text", x = anova_soas[5], y = .22, label = "*", size = 10)+ 
  scale_x_continuous(expand = c(0.02, 0), breaks = sort(unique(pdat$soa)), labels = labels) +
  scale_y_continuous(expand = c(0.05, 0)) +
  theme_classic() +
  theme(legend.position="bottom",
        axis.title = element_text(face = "bold", size = 20),
        axis.text = element_text(face = "bold", size = 12),
        legend.title = element_blank(),
        legend.text = element_text(face = "bold", size = 12)) +
  labs(x = "CTOA", y = "z-standardized RT")


tc_plot
  

if(do_save) ggsave(filename = paste0(repo_path, "tc_plot.svg"), plot = tc_plot, width = 2400, height = 1900, units = "px")

if(do_save) ggsave(filename = paste0(repo_path, "tc_plot.png"), plot = tc_plot, width = 2400, height = 1900, units = "px")

```

## Difference (invalid - valid) grand average time course

Errorbars represent Cosineau-corrected SE.

```{r difference time course ga, message = FALSE, warning=FALSE}


# RTz DIFF publication plot


pdat <- diff %>%
  summarySEwithin(data = ., 
                  measurevar = "diffz",
                  withinvars = c("soa"),
                  idvar = "subject") %>%
  mutate(soa = fac2num(soa))


labels = rep('', 25)
labels[seq(1,25,2)] = as.character(sort(unique(pdat$soa))[seq(1,25,2)])

diff_plot <- pdat %>%
  ggplot(aes(x = soa, y = diffz)) +
  geom_hline(yintercept = 0, color = "grey50", linetype = "dotted", size = 1.2)+
  geom_line(size = 1.2, alpha = .5)+
  geom_linerange(aes(min = diffz - se, 
                    max = diffz + se), size = 1.1,
                position=position_dodge(width=8))+
  geom_point(size = 2.3, position=position_dodge(width=8))+
  labs(x = "CTOA", y = "z-standardized RT difference (invalid-valid)")+
  scale_x_continuous(expand = c(0.02, 0), breaks = sort(unique(pdat$soa)), labels = labels) +
  scale_y_continuous(expand = c(0.05, 0)) +
  theme_classic() +
  theme(legend.position="bottom",
        axis.title = element_text(face = "bold", size = 20),
        axis.text = element_text(face = "bold", size = 12),
        legend.title = element_blank(),
        legend.text = element_text(face = "bold", size = 12))

diff_plot


if(do_save) ggsave(filename = paste0(repo_path, "tc_diff_plot.svg"), plot = diff_plot, width = 2400, height = 1900, units = "px")


```



# ANOVA & Post-hoc t-tests


```{r prep anova, message = FALSE, warning=FALSE}

aov_data <- clean %>%
  # subset to Posner SOAs
  filter(soa %in% anova_soas) %>%
  # convert relevant variables to factors
  mutate(validity = factor(validity),
         soa = factor(soa))

```


### Results 

Perform a repeated-measures ANOVA with z-standardized RTs as dependent variables and the CTOA and validity as within-factors. 

```{r anova res, message = FALSE, warning=FALSE, comment = ""}
# compute ANOVA (allowing GG corrections)
aov_res <- aov_data %>%
  aov_car(rtz ~ soa + validity + Error(subject/soa*validity), data = .) 


# plot ANOVA results
aov_res
```

### Post Hoc Tests

Within each family of tests (i.e. tests within each factor or interaction), multiple comparison correction was applied. Thus, p-values in the subsequent tables are already corrected.

```{r anova post hoc tests, message = FALSE, warning=FALSE}

# post hoc tests soa
contrast(emmeans(aov_res, ~ soa),
         interaction = c("pairwise"),
         adj = mcc_method) %>%
  knitr::kable(.) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))


# post hoc tests interaction
contrast(emmeans(aov_res, ~ validity + soa), 
         interaction = c("pairwise", "identity"),
         adj = mcc_method) %>%
  knitr::kable(.) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

```

